---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```


# fr

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/fr)](https://CRAN.R-project.org/package=fr)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![R-CMD-check](https://github.com/cole-brokamp/fr/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/cole-brokamp/fr/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

`fr` provides `fr_field` and `fr_tdr` objects for implementing [frictionless](https://specs.frictionlessdata.io) [tabular-data-resource](https://specs.frictionlessdata.io/tabular-data-resource) standards in R.

## Installation

You can install the development version of codec from [GitHub](https://github.com/) with:

```{r, eval = FALSE}
# install.packages("pak")
pak::pak("cole-brokamp/fr")
```
	
## Usage

```{r}
library(fr)
```

Convert R vectors to a frictionless field (i.e., `fr_field` object):

```{r}
uid <-
  replicate(34, paste0(sample(c(letters, 0:9), 8, TRUE), collapse = "")) |>
  as_fr_field(name = "uid",
              title = "Unique Identifier",
              description = "Each uid is composed of 8 random alphanumeric characters")
```

TODO Printing and other operations proceed as if `uid` were still a character:

```{r, eval = FALSE}
grep("[0-9]", uid)
```

To drop the Frictionless attributes and coerce a `fr_field` object back to a vector, use `as_vector()`:

```{r}
as_vector(uid)
```

Inspect [field descriptors](https://specs.frictionlessdata.io/table-schema/#field-descriptors) using `fr_descriptors()`:

```{r}
## fr_descriptors(uid)
```
