---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# fr

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/fr)](https://CRAN.R-project.org/package=fr)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![R-CMD-check](https://github.com/cole-brokamp/fr/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/cole-brokamp/fr/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

`fr` provides `fr_field` and `fr_tdr` objects for implementing [frictionless](https://specs.frictionlessdata.io) [tabular-data-resource](https://specs.frictionlessdata.io/tabular-data-resource) standards in R.

## Installation

You can install the development version of {fr} from [GitHub](https://github.com/) with:

```{r, eval = FALSE}
# install.packages("pak")
pak::pak("cole-brokamp/fr")
```
	
## Usage

```{r}
library(fr)
```

### Frictionless Field

Create a frictionless field (i.e., `fr_field` object) by converting an R vector and specifying named [field descriptors](https://specs.frictionlessdata.io/table-schema/#field-descriptors):

```{r}
set.seed(1)
uid <-
  replicate(6, paste0(sample(c(letters, 0:9), 8, TRUE), collapse = "")) |>
  as_fr_field(name = "uid",
              title = "Unique Identifier",
              description = "Each uid is composed of 8 random alphanumeric characters.")
```

Printing works as usual except that `fr_field` objects prepend extra lines for the Frictionless attributes:

```{r}
uid
```

Use `str()` (or `pillar::glimpse()`) to view the entire structure of the `fr_field` object:

```{r}
str(uid)
```

Extract a single descriptor using the `@` or `S7::prop` accessor functions from the {[S7](https://github.com/RConsortium/S7)} package.

```{r}
uid@title
S7::prop(uid, "description")
```

This also works to set descriptors:

```{r}

uid@title <- "Random Unique Identifier"

S7::prop(uid, "description") <- "Each identifier is randomly generated to be 8 random alphanumeric characters."

str(uid)

```

Extract just the descriptors that have a value as a list using `fr_desc`:

```{r}
fr_desc(uid)
```

`fr_field` objects can be used anywhere that the underlying vector in `@value` can be used. Because `uid`'s Frictionless `type` is `string`, the underlying vector in this case is a character vector.

```{r}
grepl("[[:alnum:]]", uid)
paste(uid, collapse = "-")
```

Explicitly drop the Frictionless attributes and extract just the underlying vector that contains the data vector with `as_vector()`:

```{r}
as_vector(uid)
```

### Frictionless Tabular-Data-Resource

Convert a data frame into a frictionless tabular-data-resource (i.e., `fr_tdr` object) with `as_fr_tdr()`.  Here, we create some metadata based on `?mtcars`:

```{r}
d_fr <-
  mtcars |>
  tibble::as_tibble() |>
  as_fr_tdr(name = "mtcars",
            version = "0.9.1",
            title = "Motor Trend Car Road Tests",
            homepage = "https://rdrr.io/r/datasets/mtcars.html",
            description = "The data was extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973–74 models).")
```

Print the `fr_tdr` object to view all of the table-specific metadata descriptors and the underlying data frame:

```{r}
d_fr
```

Use `str()` or `pillar::glimpse()` to view all field- and table-specific descriptors:

```{r}
pillar::glimpse(d_fr)
str(d_fr, max.level = 1)
```

Use `fr_desc()` to get a list of descriptors, excluding the list of `fields`:

```{r}
fr_desc(d_fr)
```

Add a metadata descriptor for one of the fields in the tabular data resource by using the `@` or `S7::prop` accessor functions from the {[S7](https://github.com/RConsortium/S7)} package:

```{r}
d_fr@fields$cyl@title <- "Number of cylinders"
S7::prop(d_fr@fields$gear, "title") <- "Number of forward gears"

d_fr@fields[c("cyl", "gear")]

## d_fr[c("cyl", "gear")]

```

Add a `name` metadata descriptor for each field in the `fr_tdr` object 

```{r}
# TODO

the_tdr <- c(fr_desc(d_fr), list(fields = lapply(d_fr@fields, fr_desc)))

str(the_tdr)

```

Using `str()` provides a useful overview of the structure of a `fr_tdr` object, including all field-specific metadata, table-specific metadata, and the underlying `@value` data vector:

```{r}
str(d_fr)
```

Use `fr_schema()` to extract the metadata for each field in a list. Pair this with [listviewer](https://cran.r-project.org/package=listviewer) for an interactive list viewer:

```{r, eval = FALSE}
fr_schema(d_fr) |>
    listviewer::jsonedit(mode = "view")
```

Accessor functions work as they do with data frames and tibbles, but return a `fr_field` or `td_tdr` object:

```{r}
d_fr[["disp"]] |> class()

d_fr$disp |> class()

d_fr["disp"] |> class()
```

`fr_field` objects can be used mostly anywhere that the underlying data frame can be used. 

```{r}
lm(mpg ~ cyl + disp + wt, data = d_fr)
```

Explicitly drop the Frictionless attributes and extract just the underlying data frame with `as_data_frame()` or `as_tibble()`:

```{r}
tibble::as_tibble(d_fr)

summary(d_fr)

summary(as_data_frame(d_fr))
```

An `fr_tdr` object is essentially a list of `fr_field`s with table-specific metadata descriptors.

Create a list of `fr_fields` using `fr_field()` and use it to create a `fr_tdr` object:

```{r}
# TODO add example for list approach
```
