---
title: "Creating a tabular-data-resource"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Creating a tabular-data-resource}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(fr)
```

To illustrate, we will create a short tibble with a lot of different types of fields. Imagine that `rating` is a factor with three possible levels (`good`, `better`, `best`), but only two of them are present in the data:

```{r}
d <-
  tibble::tibble(
    id = c("A01", "A02", "A03"),
    date = as.Date(c("2022-07-25", "2018-07-10", "2013-08-15")),
    measure = c(12.8, 13.9, 15.6),
    rating = factor(c("good", "best", "best"), levels = c("good", "better", "best")),
    ranking = c(14, 17, 19),
    impt = c(FALSE, TRUE, TRUE)
  )
```

Our example only has three rows, but in reality, any data frame imported, created, or curated using R can be used to create a tabular-data-resource.

We can see that we prepared a tibble with several different types of columns. Each column, or vector, has a native R class associated with it:

```{r}
sapply(d, class)

d
```

Convert the data frame into a `fr_tdr` object by using `as_fr_tdr()` and specifying some table-specific metadata. `as_fr_tdr()` uses the class of each column in R to automatically create all of the frictionless field-specific metadata (`name`, `type`, `constraints`).

```{r}
d_tdr <-
  d |>
  as_fr_tdr(
    name = "types_example",
    version = "0.1.0",
    title = "Example Data with Types",
    homepage = "https://geomarker.io",
    description = "This is used as an example dataset in the {fr} package vignette on `Creating a tabular-data-resource`."
  )

d_tdr
```

Print the `fr_tdr` object to view any table-specific metadata:

```{r}
d_tdr
```

Use `str()` (or `tibble::glimpse()`) to view all field- and table-specific metadata:

```{r}
str(d_tdr)
```

Reach in and update field-specific metadata:

```{r}
d_tdr <-
  d_tdr |>
  update_field("id",
               title = "Identifier",
               description = "This is a unique identifier for each study participant.")
```

It will complain if the descriptor is not valid:

```{r}
#| error: true
d_tdr |>
  update_field("id", foofy = "Identifier")
```

... or the field is not found:

```{r}
#| error: true
d_tdr |>
  update_field("foofy", title = "My foofy")
```

Write this to disk with `write_fr_tdr`:

```{r}
write_fr_tdr(d_tdr)

fs::dir_tree("types_example")
```

```{r, echo = FALSE}
fs::dir_delete("types_example")
```

